### 2023-01-30 article3.qmd
update article 3

```{r, load libraries}
library(akima)
library(broom)
library(tidyverse)
library(cmocean)
library(lubridate)
library(scales)
library(patchwork)
library(viridis)
library(ggridges)
library(ggvis)
library(pracma)
library(gbm)
library(MLmetrics)
library(purrr)
library(arrow)
library(plotly)
library(oce)
library(ggplot2)
library(latex2exp)
library(cowplot)
library(RColorBrewer)
library(vroom)
library(castr)
```

@@

Contrarily to previous works [@Estapa2013-gw; @Estapa2017-jr], we chose an empirical statistical approach instead of fixed thresholds to detect jumps. As a result, we firstly computed the slope between each point, then we computed the z-score (i.e. standard score) of each slope, that follows a normal distribution. The detection of jumps thus consisted in spotting outliers in the z-score distribution. For this, we computed the interquartile range ($IQR$) between the first ($Q_{1}$) and third ($Q_{3}$) quartiles. Positive and negative jumps were defined as, respectively, a z-score above $Q_{3} + 1.5 \times IQR$ and below $Q_{1} - 1.5 \times IQR$. We found that this method better detected outliers than the one used to despike the $c_{p}$ signal.

In order to compute $F_{small}$, we firstly computed the slope of each group of points between jumps (positive and negative, see @fig-estapa-method_article). Undefined slopes (a single point between two jumps) were removed as well as any negative slope assuming that either the so-called fit failed (leading to undetected jumps) either the transmissometer window was perturbed by something else than the pump cleaning the optical window. Afterwards, we computed the weighted average slope ($slope_{w}$) using the previously computed slopes of each remaining group in order to derive the small particles flux using the relationship of @Estapa2022

$$
F_{small} = 633(0.25~slope_{w})^{0.77}
$$

where $slope_{w}$ is multiplied by the transmissometer pathlength because @eq-estapa is an empirical POC-to-attenuance (ATN) relationship where ATN is the product of $c_{p}$ and the transmissometer pathlength.

We used the same relationship to compute $F_{big}$ by replacing $slope_{w}$ as the sum of positive jumps (negative jumps were discarded) divided by the drifting time.

@@

$F_{small}$ and $F_{big}$ were computed at each drifting depth for all floats (@fig-OST-flux). At 200 m, $F_{small}$ ranges between `r round(cflux_info_table[cflux_info_table$park_depth == '200 m',]$min_smallf, 1)` and `r round(cflux_info_table[cflux_info_table$park_depth == '200 m',]$max_smallf, 1)` mg C m$^{-2}$ day$^{-1}$ with an average of `r round(cflux_info_table[cflux_info_table$park_depth == '200 m',]$mean_smallf, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '200 m',]$std_smallf, 1)` mg C m$^{-2}$ day$^{-1}$ (average $\pm$ standard deviation). At 500 m, $F_{small}$ is comprised between `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$min_smallf, 1)` and `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$max_smallf, 1)` mg C m$^{-2}$ day$^{-1}$ with an average of `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$mean_smallf, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$std_smallf, 1)` mg C m$^{2}$ day$^{-1}$. At 1000 m, all floats measured an increase of $F_{small}$ from late May to mid-July followed by a steady decrease towards January, with $F_{small}$ ranging between `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$min_smallf, 1)` and `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$max_smallf, 1)` mg C m$^{-2}$ day$^{-1}$ and `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$mean_smallf, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$std_smallf, 1)` mg C m$^{-2}$ day$^{-1}$ on average.

@@

In this study, we focus on 14 size classes, from 102 $\mu$m to 2.5 mm. In @fig-uvp-1000m, we show the daily averaged concentrations of particles in each class size for each float at 1000 m. For the first three classes (i.e. 102-203 $\mu$m), we observe a similar pattern with concentrations starting at 4 particles/L at the end of May that reached 8 particles/L in mid-July and then decreased back to 4 particles/L towards December. For particles in the 203-406 $\mu$m range, the pattern can still be observed with a delay of 15 to 30 days between floats 6904240/4903634 and floats 6904241/4902578 which were deployed conjointly. Observed concentrations were very low at the end of May (\< 1 particle/L) then they increased by a factor of \~3-4 between mid-June and mid-July to decrease back to their initial concentrations towards December. For the remaining classes, concentrations are always below 1 particle/L with increasing uncertainty (rarer particles) in the signal. At 200 m (@fig-uvp-200), there is no clear peak following the bloom event in the smaller part of the size spectrum (\< 406 $\mu$m) whereas both floats 6904240 and 4903634 (i.e. conjointly deployed) observed the same high concentrations of particles above 406 $\mu$m at the end of May. At 500 m (@fig-uvp-500), the signal shows a slight increase in the abundance of particles in the 102-256 $\mu$m range from the end of May to the beginning of July or mid-July, depending of the float.

@@

uvp_data <- uvp_data |> pivot_longer(cols = lpm_class, names_to = 'size_class', values_to = 'conc')

# add level for order
uvp_data$size_class <- factor(uvp_data$size_class, levels = lpm_class)
uvp_data$park_depth <- factor(uvp_data$park_depth, levels = c('200 m', '500 m', '1000 m'))

# compute daily mean chla for each float at each cycle and at each drifting depth
mean_uvp_data <- uvp_data |> group_by(cycle, WMO, size_class, park_depth, juld) |> summarize(mean_conc = mean(conc, na.rm=T))
```

```{r, uvp data at 1000 m}
#| label: fig-uvp-1000m
#| fig-cap: Daily averaged concentrations of particles abundance in 14 size classes measured by the UVP6 for each float at 1000 m. Black lines and shaded areas represent, respectively, the daily mean particles abundance and the 95% confidence interval around the mean.

p1000m <- mean_uvp_data |> filter(park_depth == '1000 m') |> ggplot() + geom_point(aes(juld, mean_conc, colour = WMO, shape = WMO)) +
  geom_smooth(aes(juld, mean_conc), colour = 'black') +
  scale_color_brewer(palette = 'Dark2') + 
  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
  theme_bw() + labs(x = 'Month', y = 'Particles abundance (#/L)') +
  scale_y_continuous(trans = 'log10') +
  facet_wrap(~size_class, scales = 'free_y')

@@

```{r, uvp data at 200 m}
#| label: fig-uvp-200
#| fig-cap: Daily averaged concentrations of particles abundance in 14 size classes measured by the UVP6 for each float at 200 m. Black lines and shaded areas represent, respectively, the daily mean particles abundance and the 95% confidence interval around the mean.

p200m <- mean_uvp_data |> filter(park_depth == '200 m') |> ggplot() + geom_point(aes(juld, mean_conc, colour = WMO, shape = WMO)) +
  geom_smooth(aes(juld, mean_conc), colour = 'black') +
  scale_color_brewer(palette = 'Dark2') + 
  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
  theme_bw() + labs(x = 'Month', y = 'Particles abundance (#/L)') +
  scale_y_continuous(trans = 'log10') +
  facet_wrap(~size_class, scales = 'free_y')

```{r, uvp data at 500 m}
#| label: fig-uvp-500
#| fig-cap: Daily averaged concentrations of particles abundance in 14 size classes measured by the UVP6 for each float at 500 m. Black lines and shaded areas represent, respectively, the daily mean particles abundance and the 95% confidence interval around the mean.

p500m <- mean_uvp_data |> filter(park_depth == '500 m') |> ggplot() + geom_point(aes(juld, mean_conc, colour = WMO, shape = WMO)) +
  geom_smooth(aes(juld, mean_conc), colour = 'black') +
  scale_color_brewer(palette = 'Dark2') + 
  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
  theme_bw() + labs(x = 'Month', y = 'Particles abundance (#/L)') +
  scale_y_continuous(trans = 'log10') +
  facet_wrap(~size_class, scales = 'free_y')

@@

# # ggplot(sat_data_bbox, aes(DOY, mean_chla)) + geom_point() + geom_path()
# ggplot(sat_data_bbox, aes(DOY, q50)) + geom_point() + geom_path()

# # check float position before july
# float_position <- filter(all_floats, juld < '2022-06-15') |> drop_na(lat)
# 
# # get lat/lon range for each float for that period
# float_position |> dplyr::group_by(wmo) |> dplyr::summarise(min_lat = min(lat), max_lat = max(lat),
#                                                            min_lon = min(lon), max_lon = max(lon))
# 
# chla_sat_couple1 <- filter(chla_sat, lat > 56.5, lat < 56.85, lon > -53, lon < -52.4) |> drop_na(CHL)
# 
# tmp <- chla_sat_couple1 |> dplyr::group_by(time) |> summarise(mean_chla = mean(CHL))
# 
# ggplot(tmp, aes(x = time, y = mean_chla)) + geom_point() + geom_path()
# 
# # at deployment
# chla_sat_couple1 <- filter(chla_sat, lat > 56.7, lat < 56.9, lon > -52.5, lon < -52.3) |> drop_na(CHL)
# tmp <- chla_sat_couple1 |> dplyr::group_by(time) |> summarise(mean_chla = mean(CHL))
# ggplot(tmp, aes(x = time, y = mean_chla)) + geom_point() + geom_path()
# 
# # couple 2
# chla_sat_couple2 <- filter(chla_sat, lat > 58.9, lat < 59.1, lon > -49.4, lon < -49.2) |> drop_na(CHL)
# tmp <- chla_sat_couple2 |> dplyr::group_by(time) |> summarise(mean_chla = mean(CHL))
# ggplot(tmp, aes(x = time, y = mean_chla)) + geom_point() + geom_path()
# 
# ggplot(argo_map) + w + geom_path(aes(lon, lat, colour = WMO), size = 1) +
#   geom_point(aes(lon, lat, colour = WMO), size = 1.5) +
#   scale_color_brewer(palette = 'Dark2') +
#   theme_bw() +
#   #scale_colour_manual(values = c('#a6cee3', '#33a02c', '#b2df8a', '#1f78b4')) +
#   coord_cartesian(xlim = c(-55, -40), ylim = c(55,65)) + geom_point(data = argo_map_first, aes(lon, lat), shape = 17, size = 3) +
#   geom_point(data = argo_map_last, aes(lon, lat, colour = WMO), size = 3) +
#   geom_point(data = argo_map_last, aes(lon, lat, fill = WMO), size = 3, shape = 1) +
#   labs(x = 'longitude', y = 'latitude') +
#   #https://aosmith.rbind.io/2020/07/09/ggplot2-override-aes/#suppress-aesthetics-from-part-of-the-legend
#   guides(color = guide_legend(override.aes = list(size = c(1.5,1.5,1.5,1.5),
#                                                   shape = c(NA, NA, NA, NA)))) +
#   geom_point(data = chla_sat_couple1, aes(x = lon, y = lat), colour = "red") +
#   geom_point(data = chla_sat_couple2, aes(x = lon, y = lat), colour = "red")
#

```{r}

# all_floats_chla <- filter(all_floats, PhaseName == 'NPAR', pres <= 50, pres > 2) |> drop_na(chla) |> select(juld, pres, chla, wmo) |> mutate(datetime = as_date(juld)) #|> filter(datetime < '2022-08-01')
# 
# chla_summary <- all_floats_chla |> dplyr::group_by(wmo, datetime) |> summarize(chla_mean = mean(chla, na.rm=T))
# 
# ggplot(chla_summary, aes(x = datetime, y = chla_mean)) + geom_point() + geom_path() + facet_wrap(~wmo)

@@

>>>>

```{r, load libraries}
library(akima)
library(broom)
library(tidyverse)
library(cmocean)
library(lubridate)
library(scales)
library(patchwork)
library(viridis)
library(ggridges)
library(ggvis)
library(pracma)
library(gbm)
library(MLmetrics)
library(purrr)
library(arrow)
library(plotly)
library(oce)
library(ggplot2)
library(latex2exp)
library(cowplot)
library(RColorBrewer)
library(vroom)
library(castr)
library(reactablefmtr)
```

@@

Contrarily to previous works [@Estapa2013-gw; @Estapa2017-jr], we chose an empirical statistical approach instead of fixed thresholds to detect jumps. Therefore, we firstly computed the slope between each $c_{p}$ observation for each drifting phase. Then, we computed the z-score (i.e. standard score) of each slope, that follows a normal distribution. The detection of jumps thus consisted in spotting outliers in the z-score distribution. For this, we computed the interquartile range ($IQR$) between the first ($Q_{1}$) and third ($Q_{3}$) quartiles. Positive and negative jumps were defined as, respectively, a z-score above $Q_{3} + 1.5 \times IQR$ and below $Q_{1} - 1.5 \times IQR$. We found that this method better detected outliers than the one used to despike the $c_{p}$ signal.

In order to compute $F_{small}$, we firstly computed the slope of each group of observations between jumps (positive and negative, see @fig-estapa-method_article). Undefined slopes (a single point between two jumps) were removed as well as any negative slope assuming that either the so-called fit failed (leading to undetected jumps) either the transmissometer window was perturbed by something else than the pump cleaning the optical window. Afterwards, we computed the weighted average slope ($slope_{w}$) using the previously computed slopes of each remaining group in order to derive the small particles flux using the relationship of @Estapa2022

$$
F_{small} = 633(0.25~slope_{w})^{0.77}
$$ {#eq-estapa}

where $slope_{w}$ is multiplied by the transmissometer pathlength due to the fact that @eq-estapa is an empirical POC-to-attenuance (ATN) relationship where ATN is the product of $c_{p}$ and the transmissometer pathlength.

We used the same relationship to compute $F_{big}$ by replacing $slope_{w}$ as the sum of positive jumps divided by the drifting time.

@@

$F_{small}$ and $F_{big}$ were computed at each drifting depth for all floats (@fig-OST-flux). At 200 m, $F_{small}$ ranges between `r round(cflux_info_table[cflux_info_table$park_depth == '200 m',]$min_smallf, 1)` and `r round(cflux_info_table[cflux_info_table$park_depth == '200 m',]$max_smallf, 1)` mg C m$^{-2}$ day$^{-1}$ with an average of `r round(cflux_info_table[cflux_info_table$park_depth == '200 m',]$mean_smallf, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '200 m',]$std_smallf, 1)` mg C m$^{-2}$ day$^{-1}$ (average $\pm$ standard deviation). At 500 m, $F_{small}$ is comprised between `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$min_smallf, 1)` and `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$max_smallf, 1)` mg C m$^{-2}$ day$^{-1}$ with an average of `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$mean_smallf, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '500 m',]$std_smallf, 1)` mg C m$^{2}$ day$^{-1}$. For both drifting depths, the trend shows a decrease of $F_{small}$ from June to January. At 1000 m, all floats measured an increase of $F_{small}$ from late May to mid-July followed by a steady decrease towards January, with $F_{small}$ ranging between `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$min_smallf, 1)` and `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$max_smallf, 1)` mg C m$^{-2}$ day$^{-1}$ and `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$mean_smallf, 1)` $\pm$ `r round(cflux_info_table[cflux_info_table$park_depth == '1000 m',]$std_smallf, 1)` mg C m$^{-2}$ day$^{-1}$ on average.

@@

#### When drifting

@@

uvp_data <- uvp_data |> pivot_longer(cols = lpm_class, names_to = 'size', values_to = 'conc') |> mutate(size = size |> str_remove('NP_Size_') |> as.numeric())

# add level for order
#uvp_data$size_class <- factor(uvp_data$size_class, levels = lpm_class)
uvp_data$park_depth <- factor(uvp_data$park_depth, levels = c('200 m', '500 m', '1000 m'))

# compute daily mean chla for each float at each cycle and at each drifting depth
mean_uvp_data <- uvp_data |> group_by(cycle, WMO, size, park_depth, juld) |> summarize(mean_conc = mean(conc, na.rm=T))
```

In this study, we focus on 14 size classes, from 102 $\mu$m to 2.5 mm. In @fig-uvp-200m, @fig-uvp-500m and @fig-uvp-1000m, we show the daily averaged concentrations of particles in each class size for each float at their respective drifting depth. At 200 m, it can be observed that particles concentrations in the 102 - 512 $\mu$m range are relatively uniform between June and October before decreasing sharply towards January. Between 512 $\mu$m and 1.29 mm, concentrations are steadily decreasing from June. Above 1.29 mm, the abundance of particles is already very low (< 0.1 particle/L) in June and keeps decreasing over time.

```{r, uvp data at 200 m}
#| label: fig-uvp-200m
#| fig-cap: Daily averaged concentrations of particles abundance in 14 size classes measured by the UVP6 for each float at 200 m. Black lines and shaded areas represent, respectively, the daily mean particles abundance and the 95% confidence interval around the mean. Dashed vertical lines indicate the time at which the daily mean particles abundance is maximum.

p200m <- mean_uvp_data |> filter(park_depth == '200 m') |> ggplot() + geom_point(aes(juld, mean_conc, colour = WMO, shape = WMO)) +
  geom_smooth(aes(juld, mean_conc), colour = 'black') +
  scale_color_brewer(palette = 'Dark2') + 
  scale_x_date(labels = date_format("%m"), date_breaks = '2 month') +
  theme_bw() + labs(x = 'Month', y = 'Particles abundance (#/L)') +
  scale_y_continuous(trans = 'log10') +
  facet_wrap(~size, scales = 'free_y') #+
  #geom_vline(data = test, aes(xintercept = date_max))

tmp200m <- ggplot_build(p200m)

test <- tmp200m$data[[2]] |> filter(is.finite(y)) |> group_by(PANEL) |> summarize(index_max = which.max(y), max_date = x[index_max]) |> mutate(date_max = as_date(max_date, origin = "1970-01-01"))
test$size <- unique(uvp_data$size)

p200m <- mean_uvp_data |> filter(park_depth == '200 m') |> ggplot() + geom_point(aes(juld, mean_conc, colour = WMO, shape = WMO)) +
  geom_smooth(aes(juld, mean_conc), colour = 'black') +
  scale_color_brewer(palette = 'Dark2') + 
  scale_x_date(labels = date_format("%m"), date_breaks = '2 month') +
  theme_bw() + labs(x = 'Month', y = 'Particles abundance (#/L)') +
  scale_y_continuous(trans = 'log10') +
  facet_wrap(~size, scales = 'free_y') +
  geom_vline(data = test, aes(xintercept = date_max), linetype = 'dashed')

At 500 m, particles concentrations in the 102 - 645 $\mu$ range slowly decrease from July to October with a sharp decrease in November. Above 645 $\mu$m, the trend shows a maximum between the end of July and the first half of August though the concentrations of particles in those classes are very low (< 0.5 particle/L) on the entire timeserie.

```{r, uvp data at 500 m}
#| label: fig-uvp-500m
#| fig-cap: Daily averaged concentrations of particles abundance in 14 size classes measured by the UVP6 for each float at 500 m. Black lines and shaded areas represent, respectively, the daily mean particles abundance and the 95% confidence interval around the mean. Dashed vertical lines indicate the time at which the daily mean particles abundance is maximum.

p500m <- mean_uvp_data |> filter(park_depth == '500 m') |> ggplot() + geom_point(aes(juld, mean_conc, colour = WMO, shape = WMO)) +
  geom_smooth(aes(juld, mean_conc), colour = 'black') +
  scale_color_brewer(palette = 'Dark2') + 
  scale_x_date(labels = date_format("%m"), date_breaks = '2 month') +
  theme_bw() + labs(x = 'Month', y = 'Particles abundance (#/L)') +
  scale_y_continuous(trans = 'log10') +
  facet_wrap(~size, scales = 'free_y') #+
  #geom_vline(data = test, aes(xintercept = date_max))

tmp500m <- ggplot_build(p500m)

test <- tmp500m$data[[2]] |> filter(is.finite(y)) |> group_by(PANEL) |> summarize(index_max = which.max(y), max_date = x[index_max]) |> mutate(date_max = as_date(max_date, origin = "1970-01-01"))
test$size <- unique(uvp_data$size)

p500m <- mean_uvp_data |> filter(park_depth == '500 m') |> ggplot() + geom_point(aes(juld, mean_conc, colour = WMO, shape = WMO)) +
  geom_smooth(aes(juld, mean_conc), colour = 'black') +
  scale_color_brewer(palette = 'Dark2') + 
  scale_x_date(labels = date_format("%m"), date_breaks = '2 month') +
  theme_bw() + labs(x = 'Month', y = 'Particles abundance (#/L)') +
  scale_y_continuous(trans = 'log10') +
  facet_wrap(~size, scales = 'free_y') +
  geom_vline(data = test, aes(xintercept = date_max), linetype = 'dashed')

At 1000 m, all 4 floats measured a clear maximum in particles abundance in the 102 - 645 $\mu$m in the second half of July. Above 645 $\mu$m, the same trend can still be observed with concentrations of less than 1 particle/L.

```{r, uvp data at 1000 m}
#| label: fig-uvp-1000m
#| fig-cap: Daily averaged concentrations of particles abundance in 14 size classes measured by the UVP6 for each float at 1000 m. Black lines and shaded areas represent, respectively, the daily mean particles abundance and the 95% confidence interval around the mean. Dashed vertical lines indicate the time at which the daily mean particles abundance is maximum.

p1000m <- mean_uvp_data |> filter(park_depth == '1000 m') |> ggplot() + geom_point(aes(juld, mean_conc, colour = WMO, shape = WMO)) +
  geom_smooth(aes(juld, mean_conc), colour = 'black') +
  scale_color_brewer(palette = 'Dark2') + 
  scale_x_date(labels = date_format("%m"), date_breaks = '2 month') +
  theme_bw() + labs(x = 'Month', y = 'Particles abundance (#/L)') +
  scale_y_continuous(trans = 'log10') +
  facet_wrap(~size, scales = 'free_y') #+
  #geom_vline(data = test, aes(xintercept = date_max))

tmp1000m <- ggplot_build(p1000m)

test <- tmp1000m$data[[2]] |> filter(is.finite(y)) |> group_by(PANEL) |> summarize(index_max = which.max(y), max_date = x[index_max]) |> mutate(date_max = as_date(max_date, origin = "1970-01-01"))
test$size <- unique(uvp_data$size)

p1000m <- mean_uvp_data |> filter(park_depth == '1000 m') |> ggplot() + geom_point(aes(juld, mean_conc, colour = WMO, shape = WMO)) +
  geom_smooth(aes(juld, mean_conc), colour = 'black') +
  scale_color_brewer(palette = 'Dark2') + 
  scale_x_date(labels = date_format("%m"), date_breaks = '2 month') +
  theme_bw() + labs(x = 'Month', y = 'Particles abundance (#/L)') +
  scale_y_continuous(trans = 'log10') +
  facet_wrap(~size, scales = 'free_y') +
  geom_vline(data = test, aes(xintercept = date_max), linetype = 'dashed')

@@

#### When profiling

@fig-uvp-vertical-profiles shows that all floats measure similar patterns in the vertical profiles of particles concentrations in all size classes. In the 102-512 $\mu$m range, all 4 floats measured a local maximum between 500 and 1500 m, likely due to both sinking and remineralization processes. Above 512 $\mu$m, the relative abundance is maximum along a depth track centered on August. It can be seen among all size classes that the water column is being drained proportionally to the class size.

```{r, uvp_data on the vertical}
#| label: fig-uvp-vertical-profiles
#| fig-cap: Normalized concentrations of particles for all UVP6 vertical profiles. Plots were arranged to respect the deployment of each couple of floats (e.g. the first two rows of plots are related to co-deployed floats 6904241 and 1902578). The colorbar depicts the log$_{10}$ of normalized concentrations. Grey areas indicate that particles were not measured in the size range of interest.

particles <- all_floats |> filter(PhaseName == 'NPAR') |> select(juld, cycle, depth = pres, wmo, all_of(lpm_class), MLD) |> drop_na(NP_Size_102) |> mutate(juld = as_date(juld))

particles_bis <- particles |> pivot_longer(cols = all_of(lpm_class), names_to = 'size', values_to = 'concentration') |> mutate(size = size |> str_remove('NP_Size_') |> as.numeric())
#particles_bis$size <- factor(particles_bis$size, levels = lpm_class)

# rel_conc
particles_bis <- particles_bis |> group_by(wmo, size) |> mutate(rel_conc = (concentration-min(concentration))/(max(concentration) - min(concentration)))

# order floats by couple of deployment
particles_bis$wmo <- factor(particles_bis$wmo, levels = c(6904241,1902578,6904240,4903634))

# plot concentration
# particles_bis |> filter(depth < 200) |>
#   ggplot() + facet_grid(wmo~size) + theme_bw() +
#   geom_point(aes(x=juld, y=depth, colour=rel_conc), size=1, shape=15) +
#   #scale_colour_viridis(option = 'H', limits = c(-1,1), oob = scales::squish, na.value = 'transparent') +
#   scale_colour_viridis(option = 'H') +
#   scale_y_reverse() +
#   scale_x_date(labels = date_format("%m"), date_breaks = '2 month') +
#   labs(x = 'Month', y = 'Depth (m)') +
#   geom_path(data = particles_bis, aes(juld, MLD), colour = "white")

## plot absolute concentrations
particles_bis |> filter(depth < 2000) |>
  #group_by(size) |>
  ggplot() + facet_grid(wmo~size) + theme_bw() +
  geom_point(aes(x=juld, y=depth, colour=log10(rel_conc)), size=1, shape=15) +
  #geom_point(aes(x=juld, y=depth, colour=rel_conc), size=1, shape=15) +
  scale_colour_viridis(option = 'H', limits = c(-3,-1), oob = scales::squish, na.value = 'grey') +
  #scale_colour_viridis(option = 'H') +
  scale_y_reverse() +
  scale_x_date(labels = date_format("%m"), date_breaks = '3 month') +
  labs(colour = 'Normalized concentrations (#/L)', x = 'Month', y = 'Depth (m)') +
  #geom_path(data = particles_bis, aes(juld, MLD), colour = "black") +
  theme(legend.position = 'bottom') #+
  #geom_hline(yintercept = c(200,500,1000), linetype = 'dotted') #+
  #guides(colour = guide_legend(title.position = "bottom"))

# # interpolate to see better
# dli <- particles_bis |> mutate(juldi = as.numeric(juld)) |>
#   group_by(wmo, size) |>
#   do({
#     i <- akima::interp(x=.$juldi, y=.$depth, z=.$concentration, nx=50, ny=150, duplicate="mean") |>
#       tidy()
#     # ggplot(i) + geom_raster(aes(x=x, y=-y, fill=z))
#     rename(i, juldi=x, pres=y, concentration=z)
#   }) |>
#   # re-compute relative concentration to see stuff better
#   group_by(size) |>
#   mutate(rel_conc=concentration/max(concentration, na.rm=TRUE)) |>
#   mutate(juld = as_date(juldi, origin = "1970-01-01")) |>
#   drop_na(rel_conc) |>
#   ungroup()
# 
# dli |> 
#   ggplot() + facet_grid(wmo~size) +
#   geom_raster(aes(x=juldi, y=-pres, fill=rel_conc)) +
#   #scale_y_reverse() |>
#   scale_x_continuous(expand=c(0,0), limits=c(NA, NA)) +
#   scale_y_continuous(expand=c(0,0), limits=c(-300,0)) +
#   scale_fill_viridis_c(option = 'H')

# test_sigma <- all_floats |> filter(PhaseName == 'NPAR', sigma > 26) |> select(juld, cycle, depth = pres, wmo, sigma, MLD) |> drop_na(sigma) |> mutate(juld = as_date(juld))
# test_sigma |> ggplot() + geom_point(aes(x = juld, y = depth, colour = sigma)) + 
#   scale_colour_viridis(option = 'H', oob = scales::squish, na.value = 'transparent') +
#   scale_y_reverse(limits = c(200,0)) +
#   scale_x_date(labels = date_format("%m"), date_breaks = '2 month') +
#   labs(x = 'Month', y = 'Depth (m)') +
#   facet_wrap(~wmo, nrow = 4) + theme_bw() +
#   geom_path(data = test_sigma, aes(juld, MLD), colour = "grey")

### Remineralization processes

```{r, monthly anomaly between drifting at 500 and 1000 m}

tmp <- uvp_data |> filter(park_depth != '200 m', month > 5) |> group_by(wmo, size, month, park_depth) |> 
  summarize(monthly_mean_conc = mean(conc), monthly_median_conc = median(conc)) |>
  # compute concentration anomaly at 500 and 1000 m
  mutate(anomaly = monthly_mean_conc - lead(monthly_mean_conc)) |> drop_na(anomaly) |>
  select(wmo, size, month, anomaly) |> mutate(month = month.abb[month]) |> 
  mutate(size = as.character(size)) |>
  #mutate(anomaly = round(anomaly, 3))
  ungroup()

tt <- tmp |> filter(wmo == 6904241) |> select(size, month, anom = anomaly)
#tt <- tt |> pivot_wider(names_from = month, values_from = anom) |> mutate(size = as.character(size))

x <- c('Jun', 'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec')
y <- unique(tt$size)
data <- expand.grid(X=x, Y=y)
data$Z <- tt$anom

ggplot(data, aes(X, Y, fill= Z)) + 
  geom_tile() + scale_fill_gradientn(colors = c('darkblue', 'blue', 'white', 'red', 'darkred'))

# # https://kcuilla.github.io/reactablefmtr/articles/reactablefmtr_cookbook.html
# tt <- tt |> pivot_wider(names_from = month, values_from = anom) |> mutate(size = as.character(size))
# reactable(
#   tt,
#   pagination = FALSE,
#   showSortIcon = FALSE,
#   theme = void(
#     centered = TRUE, 
#     cell_padding = 0,
#     header_font_color = 'black',
#     font_color = 'black'
#     ),
#   defaultColDef = colDef(
#     maxWidth = 50,
#     align = 'center',
#     cell = tooltip(),
#     style = color_scales(
#       data = tt,
#       span = 2:8,
#       colors = c('#002347','#003366','#003F7D','#FF8E00','#FD7702','#FF5003'),
#       bias = 1.4,
#       opacity = 0.9,
#       show_text = FALSE
#     )
#   ),
#   columns = list(
#     total = colDef(
#       maxWidth = 225,
#       cell = data_bars(
#         data = tt,
#         fill_color = c('#002347','#003366','#003F7D','#FF8E00','#FD7702','#FF5003'),
#         bias = 1.4,
#         fill_opacity = 0.9,
#         background = 'transparent',
#         bar_height = 40,
#         text_position = 'center'
#         ),
#       style = list(borderLeft = "2px solid #999999")
#     )
#   )
# )

@@

@@

## Discussion2
